---
title: "Type I error and power simulation"
author: "Jesse Gronsbell"
date: "08/01/2022"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)
# To obtain results for this analysis please run:
# source("~/Documents/GitHub/SyntheticSurrogateAnalysis/Code/t1e_simulation.R")
# source("~/Documents/GitHub/SyntheticSurrogateAnalysis/Code/power_simulation.R")
```


## Simulation Set-up

$$
\begin{bmatrix}
           Y_i \\
           \hat{Y}_i 
         \end{bmatrix}  \mid   Z_{ik} \sim N\Big(
         \begin{bmatrix}
           \beta_G G + \beta X_i\\
           \alpha X_i
         \end{bmatrix}, \begin{bmatrix}
           1 & \rho\\
           \rho & 1\\
         \end{bmatrix}\Big)
$$

-   $G \sim Bin(2,maf)$

-   $maf = 0.25, X_i \sim N(0,1)$

-   $\alpha = \beta = 0.11$, $\beta_g = 0.11575982$ 

-   missing rate $\in\{0, 0.25 , 0.5, 0.75\}$

-   $\rho \in\{0, 0.25 , 0.5, 0.75\}$

I do not think these parameters match whats in the simulation file, please confirm.  

## Type I error 
```{r}
n.sim <- 10^4

t1e <- readRDS("~/Desktop/t1e.rds")
#"/Users/jianhuigao/Library/CloudStorage/OneDrive-UniversityofToronto/EHR_research/SyntheticSurrogateAnalysis/Data/t1e.rds")
# Fix this when you update the file.  Also follow naming convention of other files, eg. think it uses _ rather than, be consistent with variable names
# c("oracle.beta", "oracle.se", "oracle.p", "marginal.beta",
               # " marginal.se", " marginal.p", "SynSurr.beta", " SynSurr.se",
                # " SynSurr.p")

prop <- c() # call this tl1_table - in general variable name should tell you what something is
for (m in 1:4) {
  for (j in 1:4) {
    
    # Save these in your R data file - will avoid errors
    rate <- c(0, 0.25, 0.5, 0.75)[m]
    p <- c(0, 0.25, 0.5, 0.75)[j]
    
    n <- n.sim  # why not just take the length from the t1e[[m]][[j]][, 9]?
    
    # make 0.05 a parameter
    # label this data frame so its clear what is happening
    t1e.bi <- sum(t1e[[m]][[j]][, 9] < 0.05) / n.sim
    chisq.bi <- mean(t1e[[m]][[j]][, 7]^2 / t1e[[m]][[j]][, 8]^2)

    prop <- rbind(prop, round(c(rate, p, t1e.bi, chisq.bi), 2))
  }
}
prop <- prop %>% data.frame()
colnames(prop) <- c("missing", "rho", "t1e rejection", "Chisq")
prop %>% knitr::kable(booktabs = T, caption = "Proportion of test making type I error")
```

## Increased Power relative to baseline GWAS

```{r}
# calculate beta_g
cal_betag <- function(h2, maf = 0.25) {
  var <- 2 * maf * (1 - maf)
  return(sqrt(h2 / (var * (1 - h2))))
}

# Please also save chi2 for power.
result <- readRDS("~/Desktop/power.rds")

#("/Users/jianhuigao/Library/CloudStorage/OneDrive-UniversityofToronto/EHR_research/SyntheticSurrogateAnalysis/Data/power.rds")

re <- c()
for (i in 1:160) { # in general dont hard code , use length or something
  se.bi <- sapply(1:n.sim, function(j) result[[j]]$bi.se[i])
  se.target <- sapply(1:n.sim, function(j) result[[j]]$target.se[i])
  re <- c(re, mean((se.target / se.bi)^2))
}

missing_rate <- c(0, 0.25, 0.5, 0.75)
rho <- c(0, 0.25, 0.5, 0.75)
beta_g <- sapply(seq(0.001, 0.01, 0.001), function(x) cal_betag(h2 = x))
para <- expand.grid(missing_rate, rho, beta_g)
re <- cbind(para, re)
colnames(re) <- c("missing", "rho", "beta_g", "re")
re$h2 <- re$beta_g^2 * 0.375 / (re$beta_g^2 * 0.375 + 1)
```

```{r}
# Figure for efficiency analysis
color_ramp <- colorRampPalette(colors = c("white", "#0073C2FF"))
biv_palette <- color_ramp(n = 9)[c(3, 5, 7, 9)]


re %>%
  filter(h2 == 0.005) %>%
  ggplot(aes(
    x = 100*missing, y = re, group = rho,
    color = factor(rho)
  )) +
  geom_point() +
  geom_line() +
  xlab("Missing Rate (%)") +
  ylab("Relative Efficiency") +
  guides(color = guide_legend(title = expression(rho))) +
 #  ggtitle("SNP-heritability = 0.5%") +
  scale_color_manual(name = expression(rho ~ "(%)"), values = biv_palette) 
```



## Power as a function of heritability
```{r}
power <- c()
for (i in 1:160) {
  p.bi <- sapply(1:n.sim, function(j) result[[j]]$bi.p[i])
  power <- c(power, mean(p.bi < 0.05))
}
power <- cbind(para, power)
colnames(power) <- c("missing", "rho", "beta_g", "power")
power$h2 <- power$beta_g^2 * 0.375 / (power$beta_g^2 * 0.375 + 1)

power %>%
  ggplot(aes(
    x = h2*100, y = power, group = rho,
    color = factor(rho)
  )) +
  geom_point() +
  geom_line() +
  facet_wrap(.~missing)+
  xlab("SNP Heritability (%)") +
  ylab("Power") +
  guides(color = guide_legend(title = "rho")) +
  ggtitle("Power of bivariate regression, stratified by percent of missing lables")
```

Unlike the previous plot, SNP heritability plays no role in relative efficiency. 

```{r}
re %>% ggplot(aes(x = h2, y = re, group = rho, color = rho)) +
  geom_point() +
  geom_line() +
  facet_wrap(. ~ missing)
```

# Tables and Figures for the Manuscript

```{r}
# Table for main text  chi2, t1e + chi2, power for snp heritability 0.005
library(xtable)

power_table <- round(power[power[,c("h2")] == 0.005, c("missing", "rho", 
                                                       "power")], 2) # add chi2 when available
t1e_table <- prop # rename above
t1e_power_table <- inner_join(t1e_table, power_table, by = c("missing",  
                                                             "rho"))

colnames(t1e_power_table) <- c("Missing Rate (\\%)", "$\\rho$",  "Type I Error",
                               "$\\chi^2$",   "Power")  #,  "$\\chi^2$")
print(xtable(t1e_power_table), sanitize.colnames.function = identity)
```

```{r}
# Figure 1 for efficiency analysis
color_ramp <- colorRampPalette(colors = c("white", "#0073C2FF"))
biv_palette <- color_ramp(n = 9)[c(3, 5, 7, 9)]


re_plot <- re %>%
  filter(h2 == 0.005) %>%
  ggplot(aes(
    x = 100*missing, y = re, group = rho,
    color = factor(rho)
  )) +
  geom_point() +
  geom_line() +
  xlab("Missing Rate (%)") +
  ylab("Relative Efficiency") +
  guides(color = guide_legend(title = expression(rho))) +
  scale_color_manual(name = expression(rho ~ "(%)"), values = biv_palette) 

 ggsave(
      plot = re_plot,
      filename = "t1e_power_RE.png",
      device = "png",
      height = 6,
      width = 7.5,
      units = "in",
      dpi = 360
    )
    
```


```{r}
# Figure 2 for efficiency analysis

```